---
title: "Estimation of AE probabilities with savvyr"
package: savvyr
bibliography: "../inst/REFERENCES.bib"
output:
  rmarkdown::html_vignette:
          toc: true
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Estimation of AE probabilities with savvyr}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(savvyr)
library(kableExtra)
```




# Example using dummy data

We generate the dataset $S1$ in  @stegherr_meta_analytic_2021 using the parameter values for Arm A:

```{r, include=TRUE, echo=TRUE}
# sample size
N <- 200

# support of uniform censoring distribution
min.cens <- 0
max.cens <- 1000

# hazards for the three event types
haz.AE <- 0.00265
haz.death <- 0.00151
haz.soft <- 0.00227



# generate dataset
set.seed(2020)
dat1 <- generate_data(N, cens = c(min.cens, max.cens), haz.AE, haz.death, haz.soft)


# compute tau
tau <- max(dat1[, "time_to_event"])
```

The structure of the dataset looks as follows:

```{r, include=TRUE, echo=TRUE}
kable(head(dat1, 10), align = c("crcr"))
```

For this dataset we then compute all the estimators used in the comparisons in @stegherr_survival_2021 and @stegherr_estimating_2021:

```{r, include=TRUE, echo=TRUE}
# compute each estimator
IP <- inc_prop(dat1, tau)
ID <- prop_trans_inc_dens(dat1, tau)
KM <- one_minus_kaplan_meier(dat1, tau)

# competing event "death only"
IDCE2 <- prop_trans_inc_dens_ce(dat1, ce = 2, tau)
AJ2 <- aalen_johansen(dat1, ce = 2, tau)

# account for all competing events
IDCE3 <- prop_trans_inc_dens_ce(dat1, ce = 3, tau)
AJ3 <- aalen_johansen(dat1, ce = 3, tau)

# display
tab <- rbind(IP, ID, KM, IDCE2, AJ2[1:2], IDCE3, AJ3[1:2])
colnames(tab) <- c("estimated AE probability", "variance of estimation")
rownames(tab) <- c(
  "incidence proportion", "probability transform incidence density ignoring competing event",
  "1 - Kaplan-Meier", "probability transform incidence density (death only)",
  "Aalen-Johansen (death only), AE risk", "probability transform incidence density (all CEs)",
  "Aalen-Johansen (all CEs), AE risk"
)

# probability of AE
kable(tab, digits = c(3, 5))
```

Finally, the estimated probabilities of competing events based on the Aalen-Johansen estimators:

```{r, include=TRUE, echo=TRUE}
# display
tab <- rbind(AJ2[3:4], AJ3[3:4])
colnames(tab) <- c("estimated probability", "variance of estimation")
rownames(tab) <- c(
  "Aalen-Johansen (death only), CE risk",
  "Aalen-Johansen (all CEs), CE risk"
)

# probability of AE
kable(tab, digits = c(3, 5))
```


# References






